\chapter{Background theory}
\section{Position calculation with range measurement}

Position estimation with the use of range measurement apply the pseudo-range measurement, which is defined as the geometrical range in addition to all errors that has affected the measurement. A pseudo-range model can be:

\begin{equation}
y = h(x) + \beta
\end{equation}

where y is the pseudo-range, $h(x) = ||p_O - p||_2$, and $\beta$ is a stochastic disturbance in the measurement.

\section{Estimation}
Different estimation techniques that is used for position estimation. Least squares method is a simple estimation technique with good accuracy for a stationery target. However for a dynamical target a Kalman filter is more suited. Different Kalman filter, which is relevant for target tracking, and sensor fusion practise.

\section{Least square}
Apply the Moore-Penrose psudoinverse
\begin{equation}
\textbf{H}^{\dagger} = (\textbf{H}^T\textbf{W}\textbf{H})^{-1}\textbf{H}^T\textbf{W}
\end{equation}

\section{Kalman filter}
The Kalman filter was presented in the paper (cite kalman) by Rudolf Kalman, and present a filter that in the present of noise. Separate signal from noise.
Different filter that can be used. Present the linear kalman filter, EKF (Partical filter)?

\section{Sensor Fusion}
How to apply different sensors to find position.
\section{Path}
In order for a system to move autonomous from A to B a path is needed. A path can be as simple as a straight line from A to B, however either obstical in the path or restriction on the system can prohibit the exiting such a path.

A minimum requirement of a path is that it's connected. A paths connection level is described by its smoothness.
Feasbale path. Order of continuity.

\subsection{Straight lines}
The simplest for of creating a path is a straight line between two waypoints. The advantage with a straight line is that it's easy for a guidance system to follow the line, however it will experience a jump in reference when transitioning to another straight line
\subsection{Dubins path}
Dubins path is the shortest path from from one waypoint to the other which is continues.
\section{Guidance}
Different strategies that solves the flight approach problem. The problem is to ensure flight path feasibility from every direction.

First apply LOS/ILSO for a traditonal approch. 